Move Schedules: Fast persistence computations in sparse dynamic settings \thanks{\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}

### 2021-08-23 move_schedules.tex
More tweaks to SIAM format

% Sets running headers as well as PDF title and authors
\headers{An Example Article}{D. Doe, P. T. Frank, and J. E. Smith}

% Title. If the supplement option is on, then "Supplementary Material"
% is automatically inserted before the title.
\title{Move Schedules: Fast persistence computations in sparse dynamic settings
	\thanks{Submitted to the editors DATE.
		\funding{This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}
	}
}

@@

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={An Example Article},
  pdfauthor={Matt Piekenbrock, P. T. Frank, and J. E. Smith}
}
\fi

@@

 The main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$. 
 Given a  filtration $K$ with $m$ simplices and maximal dimension $d$, 
 one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration order, forming the \emph{filtration boundary matrix} $D$. 
 The reduction algorithm can  compute the persistent homology for all dimensions up to $d - 1$.
 In this case,  $D$ is a square matrix of size $m \times m$. Alternatively, a single dimension $i \leq d - 1$ can be computed, in which case one reduces a pair of matrices $D_i, D_{i+1}$, where the former has dimension $d_{i-1} \times d_i$ and the latter $d_{i} \times d_{i+1}$. If one is interested in a particular dimension of homology then the latter approach is preferred. However, the algorithmic description of the algorithms is simpler when considering the full matrix, so $D$ will always be $(m \times m)$ in the descriptions that follow. All algorithms discussed generalize to both situations. We give an example using the matrix pair $(R_1, V_1)$  below. 
\\
\\
\noindent
\textbf{Example 2.1: Reduction} Consider a triangle with vertices $u,v,w$ and edges $a = (u,w)$, $b = (v,w)$, $c = (u, v)$ whose filtration order, read as inclusions from left to right, is given as $K = (u, v, w, a, b, c)$. The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariant (2) below is satisfied.
If the $i$-th column of $R$ is nonzero, then $\mathrm{low}_R(i)$ 
is the row index of its lowest nonzero entry.
Using $\mathbb{Z}_2$ coefficients and homology in dimension 1 to simplify the presentation, the steps look at follows:
\begin{displaymath}
	\def\rowsp{0.4em}
	\def\arraystretch{0.95}
	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
	%
	\begin{blockarray}{cccc}{R_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern\rowsp  & \; 1 &    &  1 \; \topstrut \\
  		v \kern\rowsp & \;     &  1 & 1 \; \\
  		w \kern\rowsp & \; 1 &  1 &    \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; , \;
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
        a \kern\rowsp  & \; 1 &  &  \; \topstrut \\
  		b \kern\rowsp & \;  & 1 & \; \\
  		c \kern\rowsp & \;  &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\quad \xrightarrow{} \,
	\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
  		v \kern\rowsp & \;     & 1 & 1 \; \\
  		w \kern\rowsp & \; 1 &    &    \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; , \;
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		a \kern\rowsp  & \; 1 & 1 &    \; \topstrut \\
  		b \kern\rowsp & \;     & 1 &    \; \\
  		c \kern\rowsp & \;     &    & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\quad \xrightarrow{} \,
	\begin{blockarray}{cccc}{R_1}
	\kern-0.2em & b & a & c  \\
		\begin{block}{c[ccc]}
  		u \kern\rowsp  & \;     & 1  &  \; \topstrut \\
  		v \kern\rowsp &  \; 1  & 1  &  \; \\
  		w \kern\rowsp & \; 1  &     &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; , \;
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & b & a & c  \\
		\begin{block}{c[ccc]}
  		b \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
  		a \kern\rowsp & \;     & 1 & 1 \; \\
  		c \kern\rowsp & \;     &    & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\vspace*{-1em}
\end{displaymath}

The column $c$ in $R$ indicates that dimension $1$ homology is born, which in this filtration is never killed since the dimension of the complex is $2$. Similarly, the columns at $u, v, w$ in $R_0$ (not shown) are all zero, indicating three $0$-dimensional homology classes  are born, which are then killed by the pivot entries in columns $a$ and $b$ in $R_1$.

>>>>

% Sets running headers as well as PDF title and authors
\headers{Fast Persistence Computations}{M. Piekenbrock and J. Perea}

% Title. If the supplement option is on, then "Supplementary Material"
% is automatically inserted before the title.
\title{Move Schedules: Fast persistence computations in sparse dynamic settings
	\thanks{\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}
}

@@

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Fast Persistence Computations},
  pdfauthor={Matt Piekenbrock and Jose A. Perea}
}
\fi

@@

 The main output of the reduction algorithm is a matrix decomposition $R = D V$, where the persistence diagram is encoded in $R$. 
 Given a  filtration $K$ with $m$ simplices and maximal dimension $d$, 
 one assembles the elementary boundary chains $\partial(\sigma)$ as columns ordered according to the filtration order, forming the \emph{filtration boundary matrix} $D$. 
 The reduction algorithm can  compute the persistent homology for all dimensions up to $d - 1$.
 In this case,  $D$ is a square matrix of size $m \times m$. Alternatively, a single dimension $i \leq d - 1$ can be computed, in which case one reduces a pair of matrices $D_i, D_{i+1}$, where the former has dimension $d_{i-1} \times d_i$ and the latter $d_{i} \times d_{i+1}$. If one is interested in a particular dimension of homology then the latter approach is preferred. However, the algorithmic description of the algorithms is simpler when considering the full matrix, so $D$ will always be $(m \times m)$ in the descriptions that follow. All algorithms discussed generalize to both situations. We give an example using the matrix pair $(R_1, V_1)$  below. 
\textbf{Example 2.1: Reduction} Consider a triangle with vertices $u,v,w$ and edges $a = (u,w)$, $b = (v,w)$, $c = (u, v)$ whose filtration order, read as inclusions from left to right, is given as $K = (u, v, w, a, b, c)$. The reduction algorithm begins by setting $R = D$ and $V = I$, followed by left-to-right column operations until invariant (2) below is satisfied.
If the $i$-th column of $R$ is nonzero, then $\mathrm{low}_R(i)$ 
is the row index of its lowest nonzero entry.
Using $\mathbb{Z}_2$ coefficients and homology in dimension 1 to simplify the presentation, the steps look at follows:
  		u \kern\rowsp  & \; 1 &    &  1 \; \topstrut \\
  		v \kern\rowsp & \;     &  1 & 1 \; \\
  		w \kern\rowsp & \; 1 &  1 &    \; \botstrut \\
	\; , \;
        a \kern\rowsp  & \; 1 &  &  \; \topstrut \\
  		b \kern\rowsp & \;  & 1 & \; \\
	\quad \xrightarrow{} \,
  		u \kern\rowsp  & \; 1 & 1 & 1 \; \topstrut \\
  		v \kern\rowsp & \;     & 1 & 1 \; \\
	\; , \;
  		a \kern\rowsp  & \; 1 & 1 &    \; \topstrut \\
  		b \kern\rowsp & \;     & 1 &    \; \\
	\quad \xrightarrow{} \,
	\; , \;
	\vspace*{-1em}
\noindent The column $c$ in $R$ indicates that dimension $1$ homology is born, which in this filtration is never killed since the dimension of the complex is $2$. Similarly, the columns at $u, v, w$ in $R_0$ (not shown) are all zero, indicating three $0$-dimensional homology classes  are born, which are then killed by the pivot entries in columns $a$ and $b$ in $R_1$.
